\chapter{Set Theory III}

\epigraph{No one shall expel us from the paradise that Cantor has created for us.}{David Hilbert}

I wasn't able to find many good videos on this (since this is usually something covered in functions), but it's easy. I found something entertaining about infinity from \href{https://www.youtube.com/watch?v=SrU9YDoXE88}{Vsauce} though.

Sets can have a finite or infinite number of elements. Beleive it or not, we can still determine wether or not two infinite sets have the same cardinality using the following definition.

\begin{boxdefine}{Equipotent}{}
	A {\bf one-to-one correspondence} between sets $A$ and $B$ is a pairing of each object in $A$ with one and only one object in $B$, and each object in $B$ paired with one and only one object in $A$. If such a correspondence exists, than we say $A$ and $B$ are {\bf equipotent}, or have the same cardinality.
\end{boxdefine}

\begin{boxexample}{}{}
	Suppose $A=\{1,2,3,4,\dots\}$ and $B=\{1,4,9,16,\dots\}$. We can say that the one-to-one correspondence is $a=b^2$ where $a \in A, b \in B$. Thus the two sets are equipotent, and $|A|=|B|$.
\end{boxexample}

Another good definition to know is when an set is {\bf countably infinite}. Of course you can't literally count an infinite set. But it is used to distinguish smaller infinities (countable), and larger infinities (uncountable).

\begin{boxdefine}{Countable Set}{}
	An infinite set is {\bf countable} if it is equipotent to the natural numbers. 
\end{boxdefine}

\begin{boxexample}{}{}
	The odd numbers are countably infinite, and the real numbers are uncountably infinite.
\end{boxexample}

We can also determine if one infinite set is a {\bf proper subset} of another infinite set. If each member of $A$ is within $B$, but not every member of $B$ is within $A$, then $A$ is a proper subset of $B$, and $|A| < |B|$.

\begin{boxexample}{}{}
	If $A=\{2,3,4,5,\dots\}$ and $B=\{1,2,3,4,5,\dots\}$, than $A$ is a proper subset of $B$.
\end{boxexample}

\begin{boxexample}{}{}
	$\mathbb{W} \subset \mathbb{N} \subset \mathbb{Z} \subset \mathbb{Q} \subset \mathbb{R} \subset \mathbb{C}$.
\end{boxexample}

\chapter{Logic I}

\epigraph{Pure mathematics is, in its way, the poetry of logical ideas.}{Albert Einstein}

If you like watching videos, I think \href{https://www.youtube.com/watch?v=SrU9YDoXE88}{Dr. Trefor Bazett} and \href{https://www.youtube.com/watch?v=wRMC-ttjhwM}{Emily S} have great videos for this.

\begin{boxdefine}{Logical Statement}{}
A {\bf logical statement} is an expression that has a definite state of being either true or false.
\end{boxdefine}

Logic is the language of mathematics. We use logic to express (and prove!) ideas to another who speaks the language. When written, we aim to leave no doubt about our meaning. Of course, logic can also used to arrive at the correct results.

\begin{boxremark*}{}{}
	A {\bf paradox} is a statement that is both true and false, but can only exist within an inconsistant system of logic (and is not considered a statement within a consistem system of logic). For example, ``This statement is false.''
\end{boxremark*}

\begin{boxexample}{}{}
	Here are some examples of logical statements
	\begin{itemize}
		\item A triangle has three sides. (a true statement)
		\item $\sqrt 2$ is a rational number. (a false statement)
		\item $\pi + 2 \leq 5$. (a true statement)
	\end{itemize}
\end{boxexample}

\begin{boxexample}{}{}
	Here are some examples of non-logical statements
	\begin{itemize}
		\item Find the smallest natural number. (instruction, not statment)
		\item Does 3=9? (question, not statement)
		\item This statement is false. (cannot be true or false)
	\end{itemize}
\end{boxexample}

\section{Logical Connectors}

This is a list of the logical connectors. They work as you would use them in English.

\medskip
\begin{tabular}{r|c|c|l}
	\hline
	Connector & Symbol & Name & Example\\
	\hline
	Not & $\neg$ & Negative & $\neg \text{False} = \text{True}$\\
	And & $\land$ & Conjunction & $\text{True} \land \text{True} = \text{True} $\\
	Or & $\lor$ & Disjunction & $\text{True} \lor \text{False} = \text{True}$\\
	Implies (If...Then) & $\implies$ & Conditional & $\text{True} \implies \text{True} = \text{True}$\\
	Iff (If and only if) & $\iff$ & Biconditional & $\text{True} \iff \text{False} = \text{False}$\\
\hline
\end{tabular}
\medskip

\begin{boxnotation*}{}{}
	It is \emph{very} uncommon to use the set operations ``$\cup$'' for OR and ``$\cap$'' for AND, but the instructor sometimes uses these. The negative is sometimes written as ``$\sim A$'' or ``$\overline A$''.
\end{boxnotation*}

Lets define how these resolve using truth tables. The first truth table we will look at is {\bf not}:

\medskip
\begin{tabular}{c|c}
	\hline
	\multicolumn{2}{c}{NOT Truth Table}\\
	\hline
	$A$ & $\neg A$\\
	\hline
	True & False\\
	False & True\\
	\hline
\end{tabular}
\medskip

\begin{boxexample}{}{}
	If $A=\text{True}$, what is $\neg A$?
	
	$\neg A=\text{False}$.
\end{boxexample}

Then we have the AND logical connector. Again, this works how you'd expect in english.

\medskip
\begin{tabular}{c|c|c}
	\hline
	\multicolumn{3}{c}{AND Truth Table}\\
	\hline
	$A$ & $B$ & $A \land B$\\
	\hline
	True & True & True\\
	False & True & False\\
	True & False & False\\
	False & False & False\\
	\hline
\end{tabular}
\medskip

\begin{boxexample}{}{}
	If $A=\text{True}$ and $B=\text{False}$, what is $A \land B$?
	
	$A \land B = \text{False}$.
\end{boxexample}

Then the OR logical connector. This one is also works how you'd expect it in english.

\medskip
\begin{tabular}{c|c|c}
	\hline
	\multicolumn{3}{c}{OR Truth Table}\\
	\hline
	$A$ & $B$ & $A \lor B$\\
	\hline
	True & True & True\\
	False & True & True\\
	True & False & True\\
	False & False & False\\
	\hline
\end{tabular}
\medskip

\begin{boxexample}{}{}
	If $A=\text{True}$ and $B=\text{False}$, what is $A \lor B$?
	
	$A \lor B = \text{True}$.
\end{boxexample}

Implies works a little differently. You can see implies also written as ``if...then'' statements. Pay attention to how this truth table works, this one can be confusing. I'll also give you an example of why these values make sense.

\medskip
\begin{tabular}{c|c|c}
	\hline
	\multicolumn{3}{c}{Implies Truth Table}\\
	\hline
	$A$ & $B$ & $A \implies B$\\
	\hline
	True & True & True\\
	False & True & True\\
	True & False & False\\
	False & False & True\\
	\hline
\end{tabular}
\medskip

\begin{boxexample}{}{}
	Consider this example: ``If it's raining, then I will bring an unbrella.'' That means the following statements are also true:
	\begin{itemize}
		\item It is raining, so I brought an unbrella.
		\item It is not raining, but I brought an unbrella.
		\item It is not raining, so I didn't being an unbrella.
	\end{itemize}
	But, the statement ``It is raining, but I didn't bring an unbrella'', means that I am lieing.
\end{boxexample}

\begin{boxexample}{}{}
	If $A=\text{True}$ and $B=\text{False}$, what is $A \implies B$?
	
	$A \implies B = \text{False}$.
\end{boxexample}

Finally, iff (or if any only if) works simularly to above, but it works both ways.

\medskip
\begin{tabular}{c|c|c}
	\hline
	\multicolumn{3}{c}{Iff Truth Table}\\
	\hline
	$A$ & $B$ & $A \iff B$\\
	\hline
	True & True & True\\
	False & True & False\\
	True & False & False\\
	False & False & True\\
	\hline
\end{tabular}
\medskip

I will use the same example as above to illistrate the difference.

\begin{boxexample}{}{}
	Consider this example: ``I will bring an unbrella if and only if it is raining.'' That means the following statements are also true:
	\begin{itemize}
		\item It is raining, so I brought an unbrella.
		\item It is not raining, so I didn't being an unbrella.
	\end{itemize}
	But the following statements mean that I am lieing:
	\begin{itemize}
		\item It is raining, but I didn't bring an unbrella.
		\item It is not raining, but I brought an umbrella.
	\end{itemize}
\end{boxexample}

\begin{boxexample}{}{}
	If $A=\text{False}$ and $B=\text{True}$, what is $A \iff B$?
	
	$A \iff B = \text{False}$.
\end{boxexample}

{\bf Simple statements} are ``single ideas.'' They often express compairison, like ``is'', ``equals'', ``less than'', etc. They are the \emph{nouns} of logic. {\bf Logical connectors} are the \emph{verbs} of logic, and connect simple statements togeather to form {\bf compound statements}

\begin{boxdefine}{Logical Connector, Simple \& Compound Statements}{}
	A {\bf simple statement} is a ``single idea'' statement that does not use a {\bf logical connector} like ``and'', ``or'', ``not'', ``implies'', and ``iff''.
	A {\bf compound statement} contains two or more simple statements connected by {\bf logical connectors}.
\end{boxdefine}

\begin{boxexample}{}{}
	An example of a {\bf simple statement} is: $a^2=49$. An example of a {\bf compound statement} is: if $a=7$ then $a^2=49$ (also: $a=7 \implies a^2=49$). 
\end{boxexample}

Two important definitions for creating logical statements are {\bf tautologies}, and {\bf contradictions}.

\begin{boxdefine}{Tautology and Contradictions}{}
	A {\bf tautology} is a statement that is always true. A {\bf contradiction} is a statement that is always false.
\end{boxdefine}

\section{Applications of Truth Tables}

All right, so suppose we have two statements and we want to see if they're equal. One way to do this would be to try every single True/False combination and see if the statements evaluate to the same value.

\begin{boxexample}{}{}
	Are the statements $\neg (A \lor B)$ and $\neg A \land \neg B$ equal?

	\medskip
	\newcolumntype{g}{>{\columncolor{skyblue}}c}	
	\begin{tabular}{g|g|c|c|c|g|g}
		\hline
		$A$ & $B$ & $\neg A$ & $\neg B$ & $A \lor B$ & $\neg (A \lor B)$ & $\neg A \land \neg B$\\
		\hline
		F & F & T & T & F & T & T\\
		F & T & T & F & T & F & F\\
		T & F & F & T & T & F & F\\
		T & T & F & F & T & F & F\\
		\hline
	\end{tabular}
	\medskip

	And, we see that $\neg (A \lor B) = \neg A \land \neg B$.
\end{boxexample}

Allright, lets try a harder one.

\begin{boxexample}{}{}
	Are the statements $\neg P \land (Q \lor R)$ and $(\neg Q \lor P) \land (Q \lor \neg R)$ equal?

	\medskip
	\begin{tabular}{r|c c c c c c c c}
		\hline
		\rowcolor{skyblue}
		$P$ & F & T & F & T & F & T & F & T\\
		\rowcolor{skyblue}
		$Q$ & F & F & T & T & F & F & T & T\\
		\rowcolor{skyblue}
		$R$ & F & F & F & F & T & T & T & T\\
		$\neg P$ & T & F & T & F & T & F & T & F\\
		$\neg Q$ & T & T & F & F & T & T & F & F\\
		$\neg R$ & T & T & T & T & F & F & F & F\\
		$Q \lor R$ & F & F & T & T & T & T & T & T\\
		\rowcolor{skyblue}
		$\neg P \land (Q \lor R)$ & F & F & T & F & T & F & T & F\\
		$\neg Q \lor P$ & T & T & F & T & T & T & F & T\\
		$Q \lor \neg R$ & T & T & T & T & F & F & T & T\\
		\rowcolor{skyblue}
		$(\neg Q \lor P) \land (Q \lor \neg R)$ & T & T & F & T & F & F & F & T\\
		\hline
	\end{tabular}
	\medskip

	Imagine that, the two expressions I made up aren't equal!
\end{boxexample}
